---
title: "Project Overview"

knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})


output: html_document
---

```{r, echo=F, message=F, warning=F}

# initial code

#### sex overall

geschlecht <- read.csv2("data/geschlecht_cleaned.csv", stringsAsFactors = F, colClasses = "character")

geschlecht$Maennlich <- as.numeric(geschlecht$Maennlich)
geschlecht$Weiblich <- as.numeric(geschlecht$Weiblich)
geschlecht$Insgesamt <- as.numeric(geschlecht$Insgesamt)

geschlecht$male_percent <- (geschlecht$Maennlich / geschlecht$Insgesamt) * 100
geschlecht$female_percent <- (geschlecht$Weiblich / geschlecht$Insgesamt) * 100

geschlecht$difference <- geschlecht$male_percent - geschlecht$female_percent
geschlecht$difference_rounded <- round(geschlecht$difference, digits = 1)

### load fortified dataset

landkreisen_fortified <- readRDS("r-files/landkreisen_fortified")

library(dplyr)
landkreisen_fortified_geschlecht <- left_join(landkreisen_fortified, geschlecht, by = c("id" = "ID"))


```

<br><br>

### Introduction

For this project, I wanted to explore dispararities in the population of men and women across Germany. I have frequently heard about the flight of women from rural communities, especially in the former East, being a contributing factor to depression and economic inertia in these areas, so I was curious if the data would show that there are indeed rurals areas, particularly in the East, where there are significantly more men than women.

Prior to starting this project I hadn't seen any other examples of this data being analyzed in this way, though of course in the process of doing further research for it, I did eventually find other takes on this data (specifically, [this web report](https://www.iwkoeln.de/studien/iw-kurzberichte/beitrag/wido-geis-anja-katrin-orth-weniger-frauen-gehen-maennerueberschuss-bleibt-bestehen-302444.html) from the Institut der deutschen Wirtschaft). Their map will look differently from mine, because they focus on a somewhat broader age range (20 - 49-year-olds) than I do.

***

### Data Sources

My core data is population statistics for individual "Landkreisen" (administrative divisions that are at the NUTS-3 level, similar to the US "county" level) via the German statistics portal, [Destatis](https://www.destatis.de/DE/Startseite.html). Unfortunately, their site does not allow me to provide a direct link to the data source, but it can be found via the table code 12411-0016 ("Bevölkerung: Kreise, Stichtag, Geschlecht") and 12411-0018 ("Bevölkerung: Kreise, Stichtag, Geschlecht, Altersgruppen").

The shapefile I acquired from the Bundesamt für Kartographie und Geodäsie. It can be downloaded [here](http://www.geodatenzentrum.de/geodaten/gdz_rahmen.gdz_div?gdz_spr=deu&gdz_akt_zeile=5&gdz_anz_zeile=1&gdz_unt_zeile=19&gdz_user_id=0).

***

### Mapping the Disparaties ###

#### Overall Population ####

To start out with, I wanted to first get a picture of how gender disparities vary across Germany irrespective of age. The "difference" mapped here is the percentage of men living in a given Landkreis minus the percentage of women living in that same Landkreis. The difference is calculated as the percentage of a population that is male minus the percentage of the population that is female. A more negative number represents a more female-skewing Landkreis, and a more positive number represents a more male-skewing Landkreis.

<br>

*To zoom into this map, scroll with your mouse or trackpad. To bring the map back to its original size, double click within the plot*

```{r, echo=F, message=F, warning=F}
library(ggplot2)
library(viridis)
map_allages <- ggplot(data = landkreisen_fortified_geschlecht,
       aes(long, lat, fill = difference_rounded, group = group, text = paste("<b>", Kreise, "</b>", "<br>","M-F Difference: ", difference_rounded, "%", sep = ""))) +
  #scale_fill_distiller(type = "div",
  #                    palette = "RdYlBu",
  #                   name = "Percentage",
  #                  direction = -1) +
  scale_fill_viridis(option = "inferno", trans = "reverse") +
  coord_fixed() +
  geom_polygon() +
  #geom_path(lwd = 0.05) +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(size = 10)) +
  labs(title = "<b>Percent Difference between \n Male and Female Populations in Germany</b> \n All ages, by Landkreis \n",
        fill = "Percent difference \n (M minus F)")

library(plotly)

map_allages_ly <- ggplotly(map_allages, tooltip = "text")
map_allages_ly <- config(map_allages_ly, displayModeBar = F, 
                             scrollZoom = T)
style(map_allages_ly, hoveron = "fills") # "fills"" ties the hover text to the shape of the Landkreis -- stops it from grabbing data from neighboring Kreisen



```

As the scale shows (ranging from -5 percent to 2.5 percent), the overall variation in the difference between the male and female populations in a given Landkreis doesn't vary hugely. The most female-skewing Landkreis, Landau in der Pfalz in the state of Rheinland-Pfalz, has a 5 percent difference between the male and female populations. The most male-skewing Landkreis, Dingolfing-Landau in the state of Bavaria, has a 3 percent difference between the male and female populations. 

Interestingly, when looking at all ages, we don't see a clear separation between East and West. Indeed, the darkest part of the map, i.e. the area with the highest concentration of male-skewing Landkreisen, is in southern Bavaria. But all in all, these numbers aren't particularly dramatic, and the map shows that male to female rations vary noticeably across Germany. The numbers do tells us that overall Landkreisen in Germany skew slightly female, with approximately 85 percent of all Landkreisen having slightly to somewhat more women than men. 

<br>

#### Adults aged 18-35 ####

<!-- https://www.zeit.de/2016/37/landflucht-europa-provinz-vernachlaessigung -->
<!-- https://www.zeit.de/2017/17/ostdeutschland-abwanderung-frauen-bedrohung -->
<!-- https://www.zdf.de/dokumentation/37-grad/37-sag-mir-wo-die-frauen-sind-100.html-->
<!-- https://www.zdf.de/dokumentation/37-grad/37-sag-mir-wo-die-frauen-sind-100.html-->

But to me, the most interesting variation to look at is not the disparities across Germany for the population as a whole. When we break up the data into age groups, we can explore specific trends that are tied to stages of life. I wanted to specifically focus on the 18-35 bracket, because this is an age range where many people are moving to new locations for educational or work opportunities (or both, since Germany has a great apprenticeship system), as well as an age range where many people are seeking out and establishing romantic partnerships (and thus an age range where a strong imbalance in the male-female ratio can conceivably lead to social problems). How do these trends vary when looking just as this specific age bracket?

<br>

*To zoom into this map, scroll with your mouse or trackpad. To bring the map back to its original size, double click within the plot*


```{r, echo=F, message=F, warning=F}
geschlecht_age <- read.csv2("data/geschlecht_alter_cleaned.csv", 
                            stringsAsFactors = F, colClasses = "character")

geschlecht_age_1835 <- geschlecht_age[,c(1,2,8:11,25:28)]

geschlecht_age_1835[,c(3:10)] <- as.numeric(as.character(unlist(geschlecht_age_1835[,c(3:10)])))

geschlecht_age_1835$gesamt <- rowSums(geschlecht_age_1835[,3:10])

geschlecht_age_1835$male_total <- rowSums(geschlecht_age_1835[,3:6])
geschlecht_age_1835$female_total <- rowSums(geschlecht_age_1835[,7:10])

geschlecht_age_1835$male_percent <- 
  (geschlecht_age_1835$male_total / geschlecht_age_1835$gesamt) * 100

geschlecht_age_1835$female_percent <- 
  (geschlecht_age_1835$female_total / geschlecht_age_1835$gesamt) * 100

geschlecht_age_1835$difference <- 
  geschlecht_age_1835$male_percent - geschlecht_age_1835$female_percent

geschlecht_age_1835$difference_rounded <- round(geschlecht_age_1835$difference, 
                                                digits = 1)

landkreisen_fortified_age <- left_join(landkreisen_fortified, geschlecht_age_1835, by = c("id" = "ID"))

map_1835 <- ggplot(data = landkreisen_fortified_age,
       aes(long, lat, fill = difference_rounded, group = group, text = paste("<b>", Kreise, "</b>", "<br>","M-F Difference: ", difference_rounded, "%", sep = ""))) +
  scale_fill_viridis(option = "inferno", trans = "reverse") +
  coord_fixed() +
  geom_polygon() +
 # geom_path(colour="grey", lwd = 0.05) +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(size = 10)) +
  labs(title = "<b>Percent Difference between \n Male and Female Populations in Germany</b> \n 18-35 year-olds, by Landkreis \n",
       fill = "Percent difference \n (M minus F)")

map_1835_ly <- ggplotly(map_1835, tooltip = "text")
map_1835_ly <- config(map_1835_ly, displayModeBar = F, 
                             scrollZoom = T)
style(map_1835_ly, hoveron = "fills") 

# https://plot.ly/ggplot2/geom_polygon/
# https://blog.cpsievert.me/2018/01/30/learning-improving-ggplotly-geom-sf/ 

```

The first thing you might notice when you compare the map showing gender ratios for the 18-35 age bracket with the map for Germany's overall population is that this is a much darker map, suggesting there are both stronger and more numerous male-skewing regions in this age bracket. You see this reflected in the averages of the two datasets -- for the overall dataset, the average Landkreis is **49.4 percent male**. In the 18-35 dataset, the average Landkreis is **52.5 percent male**. 

However, I didn't like how this map displayed the overall variation across Germany -- I felt like the continuous color scale made it hard to see just how much or how sharply one Landkreis might differ from another. To address this, I made a second map for the 18-35 age bracket, this one with a discrete color scale where I could manually set the breaks in the scale. In this map, each consecutive darker shade represents an increase of 4 percentage points.

The choice to manually code breaks into the scale comes with the risk of critique that I'm simply manipulating the data until it shows what I want it to. I personally find it defensible, since I think it helps highlight the more interesting parts of the map without presenting a completely false picture, but it's understandble if others disagree (and that's why I thought it was also important to first include how the map looks when viewed with a continuous scale).

<br>
*To zoom into this map, scroll with your mouse or trackpad. To bring the map back to its original size, double click within the plot*


```{r, echo=F, message=F, warning=F}
landkreisen_fortified_age$brks <- cut(landkreisen_fortified_age$difference, 
                                      breaks = 
                                        c(-10, -6, -2, 2, 6, 10, 14), 
                                      labels = c("-10% to -6%",
                                "-6% to -2%", 
                                "-2% to 2%",
                                "2% to 6%",
                                "6% to 10%",
                                "10% to 14%"))

map_1835_breaks <- ggplot(data = landkreisen_fortified_age,
       aes(long, lat, fill = brks, group = group, text = paste("<b>", Kreise, "</b>", "<br>","M-F Difference: ", difference_rounded, "%", sep = ""))) +
  scale_fill_viridis(option = "inferno", 
                     direction = -1, 
                     discrete = T) +
  coord_fixed() +
  geom_polygon() +
 # geom_path(colour="grey", lwd = 0.05) +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        legend.title = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(size = 10)) + 
  labs(title = "<b>Percent Difference between \n Male and Female Populations in Germany</b> \n 18-35 year-olds, by Landkreis \n") 


map_1835_breaks_ly <- ggplotly(map_1835_breaks, tooltip = "text")
map_1835_breaks_ly <- style(map_1835_breaks_ly, hoveron = "fills")
map_1835_breaks_ly <- add_annotations(map_1835_breaks_ly,
                                      text="Percent difference \n (M minus F)",
                                      xref="paper", yref="paper",
                                      x=1.02, xanchor="left",
                                      y=.7, yanchor="bottom",
                                      legendtitle=TRUE, showarrow=FALSE)
map_1835_breaks_ly <- config(map_1835_breaks_ly, displayModeBar = F, 
                             scrollZoom = T) # suppress plotly tool bar, but allow for zooming via scrolling
layout(map_1835_breaks_ly, legend = list(
   orientation ="v",
   y = .5)
   )

# https://stackoverflow.com/questions/38733403/edit-labels-in-tooltip-for-plotly-maps-using-ggplot2-in-r
# https://stackoverflow.com/questions/38034308/legend-title-in-plotly
# https://www.rdocumentation.org/packages/plotly/versions/2.0.16/topics/config
```

Here the most extreme cases -- i.e., the most male- and female-skewing Landkreisen -- are easier to discern. In particular, it's easier to see that much of the former East Germany is sharply male-skewing, including multiple Landkreisen with the male population exceeding the female population by 10 pecent or even more. However, this is definitely not a trend that is solely present in East Germany: most of the Landkreisen in Germany have slightly to moderately more men than women, and Southern Germany especially displays several Landkreisen with at least 6 percent more men than women.

Let's look at a few of the more striking cases.

<br>

### Diving into specific cases ###

#### Whither women? The cities! ####

You'll notice several small bright spots on the maps above, generally representing areas where there are (slightly) more women than men. They seem small in number and size, but in reality they represent large numbers of people, because these bright spots are usually Germany's largest cities. Here's a table of all of the Landkreisen that have a negative population difference between men and women within the 18-35 age bracket, meaning they have more women than men:

<br>


```{r, echo=F, message=F, warning=F}
geschlecht_age_1835_women <- filter(geschlecht_age_1835, difference < 0)
geschlecht_age_1835_women <- geschlecht_age_1835_women[,c(2,12,13,14,15,16)]
geschlecht_age_1835_women <- geschlecht_age_1835_women[order(geschlecht_age_1835_women$difference),]


library(kableExtra)

kable(geschlecht_age_1835_women, row.names = F) %>% kable_styling() %>% scroll_box(width = "100%", height = "400px")
```

<br>

If you know anything about Germany, you'll notice pretty quickly that there are a lot of familiar names cropping up in this list: cities like **Bonn**, **Düsseldorf**, **Munich**, **Cologne** and **Frankfurt**, AKA some of Germany's largest cities. Many other familiar cities (like **Berlin** or **Nuremberg**) don't show up in this table, but only because they are just barely skewing male. In general, we observe that large cities tend to either have largely balanced male-to-female ratios, or they skew slightly to moderately female. 

<br>

#### OK, but... Landau in der Pfalz?? ####

Topping the above table of female-skewing Landkreisen is **Landau in der Pfalz**, a town toward Germany's southwest with 46,000 total inhabitants. Why is such a random-seeming place the most female-skewing Landkreis by far within 18-35 age bracket?

My Technologiestiftung co-worker Ben almost immediately came up with what I think is a sound theory: Laundau is home to a branch of the Unversität Koblenz-Landau. According to Wikipedia, the main study programs there are teaching, pedagogy, psychology, social sciences, and environmental sciences -- AKA a lot of fields that are traditionally studied by more women than men. It would take more in-depth research to know how good a theory this is, but at first blush, I think it makes a lot of sense. 

<br>

#### And what about the men? ####

Ah, yes, the eternal question. Here is a table of the male-skewing Landkreisen: 

<br>

```{r, echo=F, message=F, warning=F}
geschlecht_age_1835_men <- filter(geschlecht_age_1835, difference > 0)
geschlecht_age_1835_men <- geschlecht_age_1835_men[,c(2,12,13,14,15,16)]
geschlecht_age_1835_men <- geschlecht_age_1835_men[order(-geschlecht_age_1835_men$difference),]


library(kableExtra)

kable(geschlecht_age_1835_men, row.names = F) %>% kable_styling() %>% scroll_box(width = "100%", height = "400px")

# https://plot.ly/ggplot2/geom_polygon/
# https://plot.ly/r/choropleth-maps/
```

<br>

Some observations about the most male-skewing Landkreisen:

* The leading Landkreis, **Kaiserslautern**, is well-known among Americans for its large American military base. Obviously American service personnel are not included in German population statistics, but I wonder if the base attracts satellite industries related to the military and defense sector that would conceivably be male-dominated? Further, Kaiserslautern is home to a technical university which presumably has more male than female students due to the fields of study it offers.
* The next four leading Landkreisen -- **Ilm**, **Aachen**, **Darmstadt**, and **Karlsruhe** -- are also well-known for their technical universities. Chances are, students in these cities will not be the least bit surprised if you tell them they are living in some of the most male-skewing areas in Germany: they already know.
* **Dingolfing-Landau**, a Landkreis in the south of Germany, is not home to a technical university. It is however located in an area home to many auto manufacturers, including a BMW factory located within the Landkreis. I'm no expert on the automobile industry, but I presume it skews male like many other engineering industries?
* Of the 17 Landkreise with at least 10% more men than women, 7 are located in east Germany.

***

### Ways to further explore this data

The analysis here is fairly superficial -- to start out with, I just wanted to try visualizing this data and to see what kinds if insights you can draw just from plotting this data on a map. Here are some ideas I have for how to explore this data further:

* It could be interesting to come up with some covariates and to do some sort of regression to have a better picture of what are determining factors for whether a given Landkreis skews male or female. Whether or not a given Landkreis is located in east or west Germany is an obvious option, but presumably the actual predictors for skewed populations are more nuanced than that (such as the presence of a university).

* This data is also available as a time series. It would be interesting to observe how gender disparities have developed in the last 20 years and if trends can be recognized (for example, women steadily leaving rural areas and moving to urban areas). Assuming that young people in general are leaving rural areas, it could also be interesting to observe how the average age in Landkreisen in Germany has changed over time. 

* I could repeat this analysis for different age groups to see how the disparities across Germany fluctuate based on age -- presumably for the older age bracket, for example, we'd see a lot more female-skewing Landkreisen, simply because women tend to live longer. 